### 2015-04-17 presentation.Rmd
Start presentation

---
title: "Distributed Algorithms for Sparse Matrix-Vector Multiplication"
author: "Eric Hare"
date: "April 28, 2015"
output: ioslides_presentation
---

## Outline

## Introduction

## Statistics Background
In matrix/vector notation, we can define a typical linear model in statistics as:

$Y = X\beta + \epsilon$

Where:

- Y is an $n \times 1$ response vector
- X is an $n \times k$ design matrix
- $\beta$ is a $k \times 1$ parameter vector
- $\epsilon$ is an $n \times 1$ error vector (that is, $\epsilon \sim MVN(0, \Sigma)$)

## A Simple Example
Suppose we have the data given in the following table:
```{r, echo=FALSE, results='asis'}
library(ggplot2)
library(xtable)

print(xtable(mpg[1:5,c("cty", "hwy")], digits = 0), comment = FALSE, type = "html")
```

We wish to use the highway mpg to predict the city mpg.

## A Simple Example (Continued)
Then we have:

$Y = \begin{bmatrix}
        18 & 21 & 20 & 21 & 16
     \end{bmatrix}^{T}$

$X = \begin{bmatrix}
        1 & 29 \\
        1 & 29 \\
        1 & 31 \\
        1 & 30 \\
        1 & 26
     \end{bmatrix}$

$\beta = \begin{bmatrix}
        \beta_0 \\
        \beta_1
     \end{bmatrix}$

$\epsilon = \begin{bmatrix}
        \epsilon_1 & \epsilon_2 & \epsilon_3 & \epsilon_4 & \epsilon_5
     \end{bmatrix}^{T}$