### 2015-04-22 presentation.Rmd

---
title: "Distributed Algorithms for Sparse Matrix-Vector Multiplication"
author: "Eric Hare"
date: "April 28, 2015"
output: ioslides_presentation
---

## Outline

## Introduction

## Statistics Background
In matrix/vector notation, we can define a typical linear model in statistics as:

$y = X\beta + \epsilon$

Where:

- y is an $n \times 1$ response vector
- X is an $n \times k$ design matrix
- $\beta$ is a $k \times 1$ parameter vector
- $\epsilon$ is an $n \times 1$ error vector (that is, $\epsilon \sim MVN(0, \Sigma)$)

## Least Squares Estimator
The "Least Squares Estimator" of the parameter vector $\beta$ is given by $b = (X^TX)^{-1}X^Ty$

## A Simple Example
Suppose we have the data given in the following table:
```{r, echo=FALSE, results='asis'}
library(ggplot2)
library(xtable)

print(xtable(mpg[1:5,c("cty", "hwy")], digits = 0), comment = FALSE, type = "html")
```

We wish to use the highway mpg to predict the city mpg.

## A Simple Example (Continued)
Then we have:

$y = \begin{bmatrix}
        18 & 21 & 20 & 21 & 16
     \end{bmatrix}^{T}$

$X = \begin{bmatrix}
        1 & 29 \\
        1 & 29 \\
        1 & 31 \\
        1 & 30 \\
        1 & 26
     \end{bmatrix}$

$\beta = \begin{bmatrix}
        \beta_0 \\
        \beta_1
     \end{bmatrix}$

$\epsilon = \begin{bmatrix}
        \epsilon_1 & \epsilon_2 & \epsilon_3 & \epsilon_4 & \epsilon_5
     \end{bmatrix}^{T}$

## A Simple Example (Continued)
$(X^TX)^{-1} = \begin{bmatrix}
        60.2714 & -2.0714 \\
        -2.0714 & 0.0714 \\
     \end{bmatrix}$

$X^Ty = \begin{bmatrix}
        96 \\
        2797 \\
     \end{bmatrix}$

$b = (X^TX)^{-1}X^Ty = \begin{bmatrix}
        -7.7286 \\
        0.9286 \\
     \end{bmatrix}$

Hence, the equation for the line of best fit is given by $y = -7.7286 + 0.9286x$.

## SpMV Calculations
Note one of the key operations here is multiplying $X^T$ by $y$. In this example, it was almost trivial, but in general, the matrix $X$ can be very large and sparse.

In particular, there is a column in $X$ for every parameter in the parameter vector.

## R Simulation
```{r, eval=FALSE}
wearables <- read.csv("data/Example_WearableComputing_weight_lifting_exercises_biceps_curl_variations.csv")
X <- apply(wearables[,-1], 2, as.numeric)
y <- as.numeric(wearables[,1])

biggerX <- do.call("rbind", replicate(50, X, simplify = FALSE))
biggery <- rep(y, times = 50)

t(X) %*% y
t(biggerX) %*% biggery
```

## Limitations of the R (and naive) implementation
- No explicit parallelism (only thread level) by default
- Even if it was explicitly parallel, no load balancing or processor communication
- Choice of multi-core or multi-machine parallelism
- Poor cache locality

## Load Balanced Sparse Matrix-Multiplication
Nastea, Frieder, and El Ghazawi (1997) published one of the earliest proposed solutions to this issue.

## LBSM-M Overview
- $Y_i = AX_i$ where A is the sparse matrix, and $X_i$ is a sequence of dense vectors
- The size of the sequence of $X_i$ vectors is very large and not a priori known
- The resulting $Y_i$ vectors are generated and transmitted on an individual basis

## Load Balancing 
Greedy allocation for sparse pattern matrices:

$F = max_i\{\sum_{j=1}^M(nZ_{i_j})\}$

$i = 1, 2, ..., P$ represents the processors.

$i_j = \{i_1, i_2, ..., i_M\}$ represents indices of rows assigned to processor i

$nZ_{i_j}$ is the number of non-zero elements in these rows.

The goal is to minimize the function F, which is minimizing the **largest bucket size that yields the highest computing time**.

## Highly Skewed Data
In a matrix that is both highly skewed and highly sparse, splitting rows that have a significantly larger number of non-zero elements is necessary. This incurs a slight overhead and hence is only recommended for more extreme cases.

## Experimental Results
Synthetically generated data:

$f(i) = \frac{C}{i^{1 - \theta}} \text{ for } i = \{1, 2, ..., N\}$

$C = \frac{1}{\sum_{i=1}^N\frac{1}{i^{1 - \theta}}}$

$\theta \in [0, 1]$

This is a discrete distribution called the **Zipf Distribution**

## Experimental Results (Continued)
<img src="images/im1.png" width="800px">

## Experimental Results (Continued)
<img src="images/im2.png" width="800px" height="500px">

## Conclusion

## Second Paper
An architecture-aware technique for optimizing sparse matrix-vector multiplication on GPUs by Maggioni and Berger-Wolf

## Sparse Matrix Compression
<img src="images/im3.png" width="800px">

## Cache Heuristic

## Cache Heuristic (Continued)

## Results

## Results (Continued)

## Conclusion

## Third Paper

## The BRC (Blocked Row-Column) Format

## Model Parameters

## Results

## Results (Continued)

## Multiple GPUs

## Caveat: Pre-Processing Overhead

## Conclusion